---
title: "D2OE_Demo_v2"
author: "Matt Bailey"
date: "July 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This is a brief demostration of analysis performed to examine the effects of Dopamine D2 receptor Overexpression on Goal-Directed Behavior in Mice

This analysis explores behavioral differences between 2 different kinds of mice working for food rewards by making lever presses, and finds differences in how they make cost benefit choices based on different effort requirements

# Importing data sets to be used in this analysis
```{r}
# Data Frame for Worktimes and Latencies in Different Press Requirements
df_pse <- read.csv("https://raw.githubusercontent.com/mrbailey450/D2R-OE-Demo-Analysis/master/PSE_Group%20-%20DATA.csv", header = TRUE)

# Data Frame for Proportion Hold Choices in Choice Experiment
df <- read.csv("https://raw.githubusercontent.com/mrbailey450/D2R-OE-Demo-Analysis/master/D2OE_CEC-prop.csv",header = TRUE)

# Data Frame for Point of Subjective Equalities from Choice Experiment
df_temp <- read.csv("https://raw.githubusercontent.com/mrbailey450/D2R-OE-Demo-Analysis/master/D2OE_CEC_avg_Presstime_Lat.csv",header = TRUE)

```

# Loading packages to be used

```{r, message = FALSE, warning = FALSE}
# Using ggplot2 package for generating plots
library(ggplot2)

# Using dplyr for summarising means and grouping varibales in df's
library(dplyr)
```


# Examining genotype differences in behavior when working for rewards

Here, I am examining data from an experimnet simply increasing the number of lever presses mice must make to get a milk reward over days

```{r}

# Define a SEM function
sem<-function(x){
  sqrt(var(x, na.rm=T)/length(x))
}

# Quick look at the data frame
head(df_temp)

# Selecting only Chow fed subjects and sessions up to 40 presses
df_temp_chow <- subset(df_temp, Dox == "Chow")
df_temp_chow_40 <- subset(df_temp_chow, Only_40 == 1)

# Determining the genotype averages for Worktimes and Latencies
df_temp_chow_40_avg <- df_temp_chow_40 %>% 
  group_by(Press_Ratio, Genotype) %>%
  summarise(avg_Worktime = mean(na.omit(Press_Worktimes)), sem_Worktime = sem(Press_Worktimes),
            avg_Latency = mean(na.omit(Press_Latency)), sem_Latency = sem(Press_Latency))

# Plotting the worktimes as a function of the press requirment
pd <- position_dodge(0.1)
b <- ggplot(data = df_temp_chow_40_avg, aes(x = Press_Ratio, y = avg_Worktime, color = factor(Genotype)))
b +  geom_errorbar(aes(ymin = avg_Worktime+sem_Worktime, ymax = avg_Worktime-sem_Worktime), size = 1, width=.1,position=pd)+ 
geom_line(size = 2)+
geom_point(size = 3)+
  labs(x = "Press Requirement")+
  labs(y = "Time to Complete Presses (s)")+
  labs(title = "Press Requirement Effects on Work Times")

# Repeated Measures ANOVA on Worktimes
with(df_temp_chow_40, summary(aov(Press_Worktimes ~ factor(Press_Ratio) * Genotype)))


# Plotting the latencies as a function of the press requirment
pd <- position_dodge(0.1)
b <- ggplot(data = df_temp_chow_40_avg, aes(x = Press_Ratio, y = avg_Latency, color = factor(Genotype)))
b +  geom_errorbar(aes(ymin = avg_Latency+sem_Latency, ymax = avg_Latency-sem_Latency), size = 1, width=.1,position=pd)+ 
geom_line(size = 2)+
geom_point(size = 3)+
  labs(x = "Press Requirement")+
  labs(y = "Latency to Press (s)")+
  labs(title = "Press Requirement Effects on Latencies")

# Repeated Measures ANOVA on Worktimes
with(df_temp_chow_40, summary(aov(Press_Latency ~ factor(Press_Ratio) * Genotype)))


b1 <- ggplot(data = df_temp_chow_40, aes(x = Press_Latency,y = Press_Worktimes))
b1 + geom_point() +
  labs(x = "Latency to Press (s)")+
  labs(y = "Time to Complete Presses (s)")+
  labs(title = "Latency x Worktimes")

# Examining the correlation between Worktimes and Latency
with(df_temp_chow_40, cor.test(Press_Latency,Press_Worktimes)) 


# Let's examine the two groups seperatley
c <- ggplot(data = df_temp_chow_40, aes(x = Press_Latency,y = Press_Worktimes, color = Genotype))
c + geom_point() +
  labs(x = "Latency to Press (s)")+
  labs(y = "Time to Complete Presses (s)")+
  labs(title = "Latency x Worktimes for 2 Genotypes")


# Examining the correlation between Worktimes and Latency for Each Genotype
df_temp_chow_40_control <- subset(df_temp_chow_40, Genotype == "Control")
df_temp_chow_40_d2oe <- subset(df_temp_chow_40, Genotype == "D2OE")
```

# Control Mice
```{r}
with(df_temp_chow_40_control, cor.test(Press_Latency,Press_Worktimes)) 
```

# D2R-OE Mice
```{r}
with(df_temp_chow_40_d2oe, cor.test(Press_Latency,Press_Worktimes)) 

```

## Summary: Discovered 2 differences with the D2R-OE mice 
(1) Perform Work Differently - (Longer Latencies and Slower Worktimes)
(2) Relationship between Latencies and Worktimes differs - (Strongder in D2R-OE mice)


## Do these difference impact the way these mice make decisions about performing work?


To examine this I performed an experiment which gave mice a CHOICE between holding a lever down for 10seconds (Constant Work), and Pressing the lever a given number of times

```{r}

head(df)

# Getting only Chow fed mice
df_chow <- subset(df, Food == "Chow")

# Getting group averages forthe Chow fed mice
df_chow_avg <- df_chow %>% 
  group_by(Press_Req, Hold_Dur,Genotype) %>%
  summarise(avg_prop = mean(Proportion))

# Looking at the 10 second hold duration
df_chow_10 <- subset(df_chow, Hold_Dur == 10) 

# Getting group averages forthe Chow fed mice in the 10 second hold condition
df_chow_10_avg <- df_chow_10 %>% 
  group_by(Press_Req, Genotype) %>%
  summarise(avg_prop = mean(Proportion), sem_prop = sem(Proportion))

# Plotting Proportion Hold Choices in the Choice Experiment 
pd <- position_dodge(0.1)
a <- ggplot(data = df_chow_10_avg, aes(x = Press_Req, y = avg_prop, color = factor(Genotype)))
a +  geom_errorbar(aes(ymin = avg_prop+sem_prop, ymax = avg_prop-sem_prop), size = 1, width=.1,position=pd)+ 
geom_line(size = 1.5)+
geom_point(size = 3)+
  labs(x = "Press Requirement")+
  labs(y = "Proportion Hold Choices")+
  labs(title = "Effort Based Choice Behavior")

# Repeated measures ANOVA on proportion of hold choices
with(df_chow_10, summary(aov(Proportion ~ factor(Press_Req) * Genotype)))

```


